---
title: "Beach Litter Data Demonstration"
author: "Sam Dunn"
date: "August 21, 2018"
output: html_document
bibliography: Demo_bib.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
R markdown is a powerful tool for generating reports, webpages, interactive tools, etc with embedded R code.  This document is an Rmarkdown document and is intended to demonstrate some of the potential opportunities of this technology for the Alliance for the Great Lakes Beach Litter Data.  There are more robust options (namely Shiny) for full websites, and I am interested in pursuing those, but I don't have much familiarity with them yet.

We have two goals:

1. To generate an interactive web portal for finding and extracting beack litter cleanup data

2. To produce a predictive model for the type and abundance of beach litter for any given site

This document is intended to demonstrate some of the features possible for the interactive web portal and the collation and cleaning of beac litter data.

## Data Import and Cleaning
The following data were collected by a former Master's Student for her work [@Hoellein2015]  Here I will demonstrate a quick example of what sorts of interactivity is possible.
```{r Data Import}
library(readxl)
library(DT)
library(lubridate)
library(leaflet)
library(ggplot2)
library(tidyverse)
library(htmlTable)
na_beach_dat<-read_excel("Beachs_All_Data_14jan14.xlsx",sheet="North Ave")
ld_beach_dat<-read_excel("Beachs_All_Data_14jan14.xlsx",sheet="Ludington")

na_beach_dat<-na_beach_dat %>% 
  rename("Team Name"="Team Leader and Name")

beach_dat<-rbind(na_beach_dat,ld_beach_dat)
beach_dat<-beach_dat %>% 
  rename("Event_Date"="Event Date",
         "Beach"="Beach Name",
         "Team"="Team Name",
         "Weight"="Weight of Debris",
         "numVolunteers"="Number of Volunteers")


```
 We imported data from a local .csv file, but we could draw from a URL just as easily.
 
 
## Interactive Data Discovery Tables 
 Let's make a table first and then see if we need to do some cleaning.  The "datatables" function is great and is highly customizable.  First, let's subset a bit to get data we are most interested in.
 
```{r}
demo<-beach_dat[,c("Event_Date","Beach","County",
                        "Middle Latitude",
                        "Middle Longitude",
                        "Weight",
                        "numVolunteers")]

# now we need to clean up the weight column since it has both numeric and strings in it

demo$Weight<-as.numeric(gsub("Lbs","",as.character(demo$Weight)))

#lets also clean upt he date to just include the date and not the time

demo$`Event_Date`<-ymd(demo$Event_Date)




datatable(demo,
          filter='top',
          extensions='Buttons',
          options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
  
```
 
We can customize this a great deal more!  It is a R wrapper for JavaScript so anything you can do in Java we can do here too!


## Maps
Okay, let's make a map of our site. The leaflet package in R is efficient and powerful.  The popup table here is simply the means of the number of volunteers and weight of trash for the whole dataset.  This can be customized to whatever we like but requires a bit more effort.


```{r Maps!}
weight_time<-ggplot(demo, aes(x=Event_Date,y=numVolunteers,group=Beach,color=Beach))+geom_point()+theme_classic()
weight_time

ggsave(plot=weight_time, file="test.svg",height=2,width=2,units="in")

demo<-na.omit(demo)

pop_up_data<-demo %>% 
  group_by(Beach,numVolunteers,Weight,`Middle Longitude`,`Middle Latitude`) %>% 
  summarise(meanVolunteers=round(mean(numVolunteers,na.rm=T)),
            meanWeight=round(mean(Weight,na.rm=T)),
            numobs=n())



leaflet() %>% 
  addTiles() %>% 
  addMarkers(data=pop_up_data,~`Middle Longitude`,~`Middle Latitude`,
             popup=~htmlTable(data.frame(meanVolunteers,meanWeight)))
  #I inserted an html table into the map for each pin ont he map.  The formatting needs a little tweakeing but the concept works!  ~htmlTable(data.frame(Event_Date,meanVolunteers,meanWeight)
```


Another potential tool is mapview.  This is BRAND SPANKING NEW (August 16th!) and ah snot been fully de-bugged.  I gave it a try here.  It is designed to work within a relational database context which is not the same structure as what we have been using above.  That being said these beach cleanup data are perfect candidates for a relational database organization structure.
```{r}
library(mapview)
data(breweries)
mapview(breweries)

```



## References