---
title: "Demonstration"
author: "Sam Dunn"
date: "August 20, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
R markdown is a powerful tool for generating reports, webpages, interactive tools, etc with embedded R code.  This document is an Rmarkdown document and is intended to demonstrate some of the potential opportunities of this technology for the Alliance for the Great Lakes Beach Litter Data

We have two goals:

*1 To generate an interactive web portal for finding and extracting beack litter cleanup data
*2 To produce a predictive model for the type and abundance of beach litter for any given site


The following data were collected by a former Master's Student for her work [citation here]  Here I will demonstrate a quick example of what sorts of interactivity is possible.
```{r Data IMport}
library(readxl)
library(DT)
library(lubridate)
library(leaflet)
library(ggplot2)
beach_dat<-read_excel("Beachs_All_Data_14jan14.xlsx",sheet="North Ave")

```
 We imported data from a local .csv file, but we could draw from a URL just as easily.
 
 Let's make a table first and then see if we need to do some cleaning.  The "datatables" function is great and is highly customizable.  First, let's subset a bit to get data we are most interested in.
 
```{r}
north_ave<-beach_dat[,c("Event Date",
                        "Middle Latitude",
                        "Middle Longitude",
                        "Weight of Debris",
                        "Number of Volunteers")]
# now we need to clean up the weight column since it has both numeric and strings in it

north_ave$`Weight of Debris`<-as.numeric(gsub("Lbs","",as.character(north_ave$`Weight of Debris`)))

north_ave<-na.omit(north_ave)

#lets also clean upt he date to just include the date and not the time

north_ave$`Event Date`<-ymd(north_ave$`Event Date`)


datatable(north_ave,
          filter='top',
          extensions='Buttons',
          options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
  
```
 
We can customize this a great deal more!  It is a R wrapper for JavaScript so anything you can do in Java we can do here too!

Okay, let's make a map of our site. The leaflet package in R is efficient and powerful


```{r}
weight_time<-ggplot(north_ave, aes(x=`Event Date`,y=`Number of Volunteers`))+geom_point()+geom_line()
weight_time

na_popup<-paste0("<b>","Mean Volunteers ","</b/>",mean(north_ave$`Number of Volunteers`,na.rm=T),"<br>",
                 "<b>","Mean Weight of Debris ","</b/>",round(mean(north_ave$`Weight of Debris`,na.rm=T)),2)

leaflet(north_ave) %>% 
  addTiles() %>% 
  addMarkers(~`Middle Longitude`,~`Middle Latitude`,popup=~na_popup)
  
```

